---
title: "Examen_Data_visualisation"
author: "pacman"
date: "June 25, 2017"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center>
  <h3><u>Compte rendue - Examen Final ~ Data Visualisation</u></h3>
  <h4>Réalisé par : Ayoub RMIDI <br> Encadré par : Mr. Elhassouny azzedine</h4>
</center>

## Exercice 1 :

Réalise une application shiny d'exploration des donnees par fonction telles que head, summarize et autres <br/><br/>
Dans l'interface : <br/><br/>

- L'application permet a l'utilisateur de donner
    * la base a explorer
    * la fonction qui permet d'explorer les donnees

- Resultats : 
    * afficher le resultat de la fonction


```{r, echo = TRUE}
shinyApp(
  
  ui = pageWithSidebar(
      headerPanel('Execution : '),
      sidebarPanel(
        # selectInput("db", label = "choose a data set to explore",
        #             choices = c("iris", "df_pop_country"), selected = "iris")
        # give the user to choose the data set to play with
        fileInput(inputId = "file",label = "Upload a data set to explore",
                  accept = c(
                    "text/csv",
                    "text/comma-separated-values,text/plain",
                    ".csv")
        ),
        # give the user the ability to see the header content or not , by default it's true
        selectInput(inputId = "head",label = "Header",
                    choices = c("TRUE","FALSE"),selected = "TRUE"),
        # give the user the ability to choose a function to apply on the uploaded dataset
        selectInput("funct", label = "choose a function to try",
                    choices = c("head", "summarize", "tail"), selected = "head")
      ),
      # here wi fix the main panel 
      mainPanel(
        textOutput("funct"),
        textOutput("dataSet"),
        tableOutput("final"),
        verbatimTextOutput("summary"),
        tags$style(type="text/css",
                 ".shiny-output-error { visibility: hidden; }",
                 ".shiny-output-error:before { visibility: hidden; }"
        )
      )
),
  
  server = function(input, output) {
      output$final <- renderTable({
        # select the dataset from the user
        inFile <- input$file
        if (is.null(inFile))
          return(NULL)
        # select the head value 
        headerContent <- reactive({  input$head })
        # read the csv file beyond the header option
        if(headerContent() == "TRUE")
          dataSet <- read.csv(inFile$datapath, header = TRUE)
        else
          dataSet <- read.csv(inFile$datapath, header = FALSE)
        # select the function from the user
        selectedFunction <- reactive({ input$funct })
        switch(selectedFunction(),
               head = {
                 output$summary <- renderPrint({ invisible(NULL)  })
                 head(dataSet)
               },
               tail = {
                 output$summary <- renderPrint({ invisible(NULL)  })
                 tail(dataSet)
               },
               summarize = {
                 output$summary <- renderPrint({
                   summary(dataSet)
                 })
               })
      })
    },
  
  options = list(height = 500)
)
```

## Exercice 2 :

Réalise une application shiny pour afficher la matrice de correlation.<br/><br/>
Dans l'interface : <br/><br/>

- L'application permet a l'utilisateur de donner 
    * la base a explorer
    * les variables a explorer soit sous forme des checkbox ou des selectize multiple d'une facon dynamique selon le fichier csv mis en upload
      
- Resultats : 
    * affiche la matrice de correlation
         
```{r, echo = TRUE}
library(corrplot)
library(htmltools)
source("sortable.R", local = TRUE)

shinyApp(
  
  ui = pageWithSidebar(
        headerPanel('Execution 2 :'),
        sidebarPanel(
          
          # give the user to choose the data set to play with
          fileInput(inputId = "file",label = "Upload a data set to explore",
                    accept =
                      c("text/csv","text/comma-separated-values,text/plain",".csv")
          ),
          # give the user the ability to choose the file's variables with a checkbox or with a selectize
          radioButtons("variablesStyle", "Variable Selection Style", c("Checkbox", "Selectize"), inline = T),
          helpText("Choose the variables to display."),
          
          # show the checkbox if the user choose to show the variables as a checkbox
          conditionalPanel("input.variablesStyle === 'Checkbox'",
                            sortableCheckboxGroupInput("variablesCheckbox", "", 
                            c("Loading..."))),
          
          # show the selectize if the user choose to show the variables as a selectize
          conditionalPanel("input.variablesStyle === 'Selectize'",
                            sortableSelectizeInput("variables", "", 
                            c("Loading..."),
                            multiple = T, options = list(plugins = 
                            list("remove_button"))))
        ),
        # here wi fix the main panel 
        mainPanel(
          plotOutput("corrPlot"),
          tags$style(type="text/css",
                 ".shiny-output-error { visibility: hidden; }",
                 ".shiny-output-error:before { visibility: hidden; }"
          )
        )
),
  
  server = function(input, output, session) {
      # read the input file uploaded by the user
      dataset <- reactive({
        inFile <- input$file
        if(is.null(inFile)) {
          NULL
        } else {
          #TODO: Better way to unescape e.g. \\t
          #read.delim(inFile$datapath, sep = gsub("\\t", "\t", input$datafile_sep, fixed
                     # = TRUE))
          read.csv(inFile$datapath)
        }
      })
      # End of dataset
      
      # get the header names of the uploaded file with a numeric format
      numericColumns <- reactive({
        df <- dataset()
        colnames(df)[sapply(df, is.numeric)]
      })
      # End of numericColumns
      
      # draw the correlated matrix in analogie with the selected variables
      correlation <- reactive({
        data <- dataset()
        variables <- input$variables
        if(is.null(data) ) {
          NULL
        } else {
          cor(dataset()[,input$variables])
        }
      })
      # End of correlation
      
      # Update variable selection
      observe({
        updateCheckboxGroupInput(session, "variablesCheckbox", choices = 
                                   numericColumns(), selected = numericColumns())
        updateSelectInput(session, "variables", choices = numericColumns(), selected = 
                            numericColumns())
      })
      
      # Link Variable Selection
      observe({
        if(input$variablesStyle == "Checkbox") {
          updateCheckboxGroupInput(session, "variablesCheckbox", selected = 
                                   isolate(input$vairables))
        }
      })
      
      observe({
        updateSelectInput(session, "variables", selected = input$variablesCheckbox)
      })
      
      # Correlation Plot 
      output$corrPlot <- renderPlot({
        val <- correlation()
        if(is.null(val)) return(NULL)

        val[is.na(val)] <- 0
        args <- list(val, order = "original", plotCI = "n")
    
        do.call(corrplot, c(list(method = "circle", type = "full"), args))
      })
      
    },
  
  options = list(height = 500)
)
```

## Exercice 3 :

Réalise une application shiny d'exploration des donnees par graphique tels que l'histogramme, scatterplot et autres. <br/><br/>
Dans l'interface : <br/><br/>

- L'application permet a l'utilisateur de donner 
    * la base a explorer
    * les types de graphiques avec les variables .
      
- Resultats : 
    * affiche le graphique ou les graphiques

```{r, echo = TRUE}
library(ggplot2)
library(scatterplot3d)
shinyApp(
  
  ui = pageWithSidebar(
      headerPanel('Execution : '),
      sidebarPanel(
        # selectInput("db", label = "choose a data set to explore",
        #             choices = c("iris", "df_pop_country"), selected = "iris")
        # give the user to choose the data set to play with
        fileInput(inputId = "file",label = "Upload a data set to explore",
                  accept = c(
                    "text/csv",
                    "text/comma-separated-values,text/plain",
                    ".csv")
        ),
        # give the user the ability to choose a function to apply on the uploaded dataset
        selectInput("funct", label = "choose any type of histogram",
                    choices = c("hist", "ggplot", "scatterplot3d"), selected = "hist")
      ),
      # here wi fix the main panel 
      mainPanel(
        tableOutput("final"),
        plotOutput("dataset"),
        tags$style(type="text/css",
                 ".shiny-output-error { visibility: hidden; }",
                 ".shiny-output-error:before { visibility: hidden; }"
        )
      )
),
  
  server = function(input, output) {
      output$final <- renderTable({
        # select the dataset from the user
        inFile <- input$file
        if (is.null(inFile))
          return(NULL)

        # read the csv file beyond the header option
        dataSet <- read.csv(inFile$datapath, header = TRUE)
        # select the function from the user
        selectedFunction <- reactive({ input$funct })
        #get the length of variables
        #my_vars <- length(names(dataSet))
        switch(selectedFunction(),
               hist = {
                 output$dataset <- renderPlot({ hist(dataSet[,3])  })
               },
               ggplot = {
                 output$dataset <- renderPlot({ 
                      ggplot(dataSet,aes(x= dataSet[,4],y=dataSet[,3]) ) + geom_point()
                   })
               },
               scatterplot3d = {
                 output$dataset <- renderPlot({
                      scatterplot3d(x= dataSet[,5],y=dataSet[,3],z=dataSet[,4],
                      color = 'blue',
                      grid=TRUE)
                 })
               })
      })
    },
  
  options = list(height = 500)
)
```
